Comprehensive Quality-Aware Automated Semantic Web Service Composition

### 2017-05-11 paper.tex
Camera-ready v1

In this paper we are only interested in robust compositions where only $exact$ and $plugin$ matches are considered, see \cite{lecue2009optimizing}. As argued in \cite{lecue2009optimizing} $plugin$ matches are less preferable than $exact$ matches due to the overheads associated with data processing. We suggest to consider the semantic similarity of concepts when comparing different $plugin$ matches. For concepts $a, b$ in $\mathcal{O}$ the \emph{semantic similarity} $sim(a, b)$ is calculated based on the edge counting method defined in Eq. (\ref{equation1}) from \cite{shet2012new}, where $N_a$, $N_b$ and $N_c$ measure the distances from concept $a$, concept $b$, and the closest common ancestor $c$ of $a$ and $b$ to the top concept of the ontology $\mathcal{O}$, respectively. 
\vspace{-0.2cm}

\begin{equation}
sim(a, b){=} \frac{2N_c \cdot e^{-\lambda L/D} }{N_{a}+N_{b}}
\label{equation1}
\end{equation}
\noindent For our purposes, $\lambda$ can be set to 0 as we do not measure the similarities of neighbourhood concepts, the matching type not considered in this paper.

Given a service request $T=(I_T,O_T)$, we represent a service composition solution for $T$ with services $S_1,\ldots,S_n$  by a weighted DAG, $WG=(V,E)$ with node set $V=\{Start, S_1, S_2, \ldots, S_n, End\}$ and edge set $E = \{e_{1}, e_{2},... e_{m} \}$. $Start$ and $End$ are two special services defined as $Start = (\emptyset, I_T, \emptyset )$ and $End  = (O_T, \emptyset, \emptyset)$ that account for the input and output requirements given by the request. Each edge $e$ from a service $S$ to a service $S'$ means that service $S$ produces an output $a\in O_S$ that is matched ($exact$ or $plugin$) to an input $b\in I_{S'}$ to be consumed by service $S'$ in the composition. Based on the matchmaking type ($type_e$), the \emph{semantic similarity} ($sim_e$) of edge $e$ can be defined as follows:
\vspace{-0.2cm}
\begin{equation}
\label{equation10}
type_e = 
\begin{cases}
	1 & \text{ if $a\equiv b$ ($exact$ match)},\\
	p & \text{ if $a \sqsubseteq b$ ($plugin$ match)}
\end{cases}
\end{equation}
\begin{equation}
\label{equation11}
sim_e = sim(a,b) = \frac{2N_c}{N_{a}+N_{b}}
\end{equation}

@@

The \emph{semantic matchmaking quality} of the service composition can be obtained by aggregating $type_e$ and $sim_e$, respectively, over all edges in $E$ as follows:
\vspace{-0.2cm}
\begin{equation}
\label{equation6}
MT {=} \prod_{j=1}^{m} type_ {e_{j}}
\end{equation}
\begin{equation}
\label{equation7}
SIM {=} \frac{1}{m}\sum_{j=1}^m sim_ {e_{j}}
\end{equation}
%\begin{equation}
%\label{eqCQ}
%CQ \stackrel{.}{=} (MT, \  SIM, \  A,\  R,\  C,\  T)
%\end{equation}
The \emph{QoS} of the service composition can be obtained by aggregating the QoS values of the participating services \cite{ma2015hybrid}. For a service composition with services $ S_1, S_2, ... S_n$ we obtain the reliability $R=\prod\limits^n_{k=1}r_{S_k}$, the availability $A=\prod\limits^n_{k=1}a_{S_k}$, the cost $C=\sum\limits^n_{k=1}c_{S_k}$, and the response time $T$ is the time of most time-consuming path in the composition, i.e., $$T = MAX \{\sum\limits^{\ell_j}_{k=1}t_{S_k} | \text{ $j\in\{1,\ldots,m\}$ and $\ell_j$ is the length of a path $P_j$}\}.$$

%Consequently, in Eq. \ref{eqCQ}, the comprehensive quality $CQ$ consists of the aggregated semantic matchmaking quality $SM$ and the QoS calculation in DAG discussed before.
\vspace{-0.1cm}
\noindent When multiple quality criteria are involved into decision making, then the overall fitness of a solution can be defined as a weighted sum of the individual criteria in Eq \ref{equation8}, assuming the preference of each quality criterion is known by users. 
\vspace{-0.2cm}
\begin{equation}
\label{equation8}
Fitness = w_1 \hat{MT} + w_2 \hat{SIM} + w_3 \hat{A} + w_4 \hat{R} + w_5(1 - \hat{T}) + w_6(1 - \hat{C})
\end{equation}
\noindent with $\sum_{k=1}^{6} w_k= 1$. We call this objective function the \emph{comprehensive quality model} for service composition.
The weights can be adjusted according to users' preferences. $\hat{MT}$, $\hat{SIM}$, $\hat{A}$, $\hat{R}$, $\hat{T}$, and $\hat{C}$ are normalised values calculated within the range from 0 to 1 using Eq. (\ref{equation9}). To simplify the presentation we also use the notation $(Q_1,Q_2,Q_3,Q_4,Q_5,Q_6) = (MT,SIM,A,R,T,C)$. $Q_1$ and $Q_2$ have minimum value 0 and maximum value 1. The minimum and maximum value of $Q_3$, $Q_4$, $Q_5$, and $Q_6$ are calculated across all task-related candidates in the service repository $\mathcal{SR}$ using the greedy search in \cite{ma2015hybrid,da2016genetic}.

\begin{equation}
\label{equation9}
\hat{Q_k} = 
\begin{cases}
	\frac{Q_k - Q_{k, min}}{Q_{k, max} - Q_{k, min}} & \text{ if $k=1,\ldots,4$ and }Q_{k, max} - Q_{k, min} \neq 0,\\
	\frac{Q_{k,max} - Q_k}{Q_{k, max} - Q_{k, min}} & \text{ if $k=5,6$ and }Q_{k, max} - Q_{k, min} \neq 0,\\
	1 & \text{ otherwise}.
\end{cases}
\end{equation}

\noindent To solve the composition task satisfactorily our goal is to maximize the objective function in Eq. (\ref{equation8}).

@@

\subsection{The Algorithms for our PSO-based Approach}\label{POS-based_algomargin}
The overall algorithm investigated here is made up of a PSO-based web service composition technique (Algorithm \ref{novelSteps}) and a $WG$ creating technique from a service queue (Algorithm \ref{graph_building}). In Algorithm \ref{novelSteps}, the  steps $4$, $5$, $6$ and $7$ are different from those of standard PSO: In step 4, the size of task-related service candidates generated by a greedy search determines the size of each particle's position. Each service candidate in a created service candidates queue is mapped to an index of a particle’s position vectors, where each vector has a weight value between 0.0 and 1.0. In step 5, service candidates in the queue are sorted according to their corresponding weight values in descending order. In step 6, this sorted queue is used as one of the inputs of the forward decoding Algorithm \ref{graph_building} to create a $WG$. In step 7, the fitness value of the created $WG$ is the fitness value of the particle calculated by the comprehensive model discussed in Section \ref{Comprehensive_Quality_Model}.
\vspace{-0.5cm}
\begin{algorithm}
 %\LinesNumbered
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{generateWeightedGraph}{generateWeightedGraph}
 \SetKwProg{Procedure}{Procedure}{}{}
 \SetNlSty{}{}{:}
 Randomly initialise each particle in the swarm\;
  \While {max. iterations not met}{
     \ForEach{particle in the swarm}{
     Create a service candidates queue and map service candidates to a particle's position vectors\;
     Sort the service queue by position vectors' weights\;
     Use Algorithm \ref{graph_building} to create a $WG$ from the service queue \;
     Calculate the $WG$ fitness value\;

@@

\begin{algorithm}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{createWeightedDAG}{createWeightedDAG}
 \SetKwProg{Procedure}{Procedure}{}{}
 %\LinesNumbered
 \SetNlSty{}{}{:}
 % \Procedure{}{
 \Input{ $I_T$, $O_T$, $queue$, $cache$}
 \Output{WG}
 $WG = (V, E)$\;
 $V \leftarrow$ \{$Start$, $End$ \} \;
 $OutputSet \leftarrow$ \{$I_{T}$\}\;
  \While { $O_{T}$ not satisfied by $OutputSet$}{
     \ForEach{$S$ in $queue$}{
      \uIf{$I_{S}$ satisfied by $OutputSet$}{  
        insert $S$ into $V$ \;  
        adjoin $O_{S}$ to $OutputSet$\;
        $queue$.remove $S$\;   
        $e \leftarrow$ calculate $type_e$, $sim_e$ using $cache$\;
        insert $e$ into $E$\;
       }	
     }
  }
 remove $dangling$ $nodes$ and $edges$ from $WG$\; 
 \KwRet $WG$\;
 %}
 \caption{Create a $WG$ from a sorted service queue.}
\label{graph_building}
\end{algorithm}

>>>>

In this paper we are only interested in robust compositions where only $exact$ and $plugin$ matches are considered, see \cite{lecue2009optimizing}. As argued in \cite{lecue2009optimizing} $plugin$ matches are less preferable than $exact$ matches due to the overheads associated with data processing. We suggest to consider the semantic similarity of concepts when comparing different $plugin$ matches. For concepts $a, b$ in $\mathcal{O}$ the \emph{semantic similarity} $sim(a, b)$ is calculated based on the edge counting method in a taxonomy like WorldNet or Ontology using Eq. (\ref{eq_sim}). This method \cite{shet2012new} has the advantages of simple calculation and good performance.  In Eq. (\ref{eq_sim}), $N_a$, $N_b$ and $N_c$ measure the distances from concept $a$, concept $b$, and the closest common ancestor $c$ of $a$ and $b$ to the top concept of the ontology $\mathcal{O}$, respectively. 
\begin{equation}
sim(a, b){=} \frac{2N_c \cdot e^{-\lambda L/D} }{N_{a}+N_{b}}
\label{eq_sim}
\end{equation}
\noindent For our purposes, $\lambda$ can be set to 0 as we do not measure the similarities of neighbourhood concepts, the matching type not considered in this paper.

Given a service request $T=(I_T,O_T)$, we represent a service composition solution for $T$ with services $S_1,\ldots,S_n$  by a weighted DAG, $WG=(V,E)$ with node set $V=\{Start, S_1, S_2, \ldots, S_n, End\}$ and edge set $E = \{e_{1}, e_{2},... e_{m} \}$. $Start$ and $End$ are two special services defined as $Start = (\emptyset, I_T, \emptyset )$ and $End  = (O_T, \emptyset, \emptyset)$ that account for the input and output requirements given by the request. Each edge $e$ from a service $S$ to a service $S'$ means that service $S$ produces an output $a\in O_S$ that is matched ($exact$ or $plugin$) to an input $b\in I_{S'}$ to be consumed by service $S'$ in the composition. Based on the matchmaking type ($type_e$), the \emph{semantic similarity} ($sim_e$) of edge $e$ can be defined as follow:

\begin{align*}
type_e = 
\begin{cases}
	1 & \text{ if $a\equiv b$ ($exact$ match)}\\
	p & \text{ if $a \sqsubseteq b$ ($plugin$ match)}
\end{cases}
&&
sim_e = sim(a,b) = \frac{2N_c}{N_{a}+N_{b}}
\end{align*}

%\begin{equation}
%\label{equation10}
%type_e = 
%\begin{cases}
%	1 & \text{ if $a\equiv b$ ($exact$ match)},\\
%	p & \text{ if $a \sqsubseteq b$ ($plugin$ match)}
%\end{cases}
%\end{equation}
%\begin{equation}
%\label{equation11}
%sim_e = sim(a,b) = \frac{2N_c}{N_{a}+N_{b}}
%\end{equation}

@@

The \emph{semantic matchmaking quality} of the service composition can be obtained by aggregating $type_e$ and $sim_e$, respectively, over all edges in $E$ as follow:
\begin{align*}
MT {=} \prod_{j=1}^{m} type_ {e_{j}}
&&
SIM {=} \frac{1}{m}\sum_{j=1}^m sim_ {e_{j}}
\end{align*}

%\begin{equation}
%\label{equation6}
%MT {=} \prod_{j=1}^{m} type_ {e_{j}}
%\end{equation}
%\begin{equation}
%\label{equation7}
%SIM {=} \frac{1}{m}\sum_{j=1}^m sim_ {e_{j}}
%\end{equation}
The \emph{QoS} of the service composition can be obtained by aggregating the QoS values of the participating services \cite{ma2015hybrid}. For a service composition with services $ S_1, S_2, ... S_n$ we obtain the reliability $R=\prod\limits^n_{k=1}r_{S_k}$, the availability $A=\prod\limits^n_{k=1}a_{S_k}$, the cost $C=\sum\limits^n_{k=1}c_{S_k}$, and the response time $T$ is the time of most time-consuming path in the composition, i.e., $$T = MAX \{\sum\limits^{\ell_n}_{k=1}t_{S_k} | \text{ $\ell_n$ is the length of a path $P_n$ and $n\in\{1,\ldots,n\}$}\}.$$

\noindent When multiple quality criteria are involved into decision making, then the overall fitness of a solution can be defined as a weighted sum of the individual criteria in Eq \ref{eq_fitness}, assuming the preference of each quality criterion is known by users. 
\begin{equation}
\label{eq_fitness}
Fitness = w_1 \hat{MT} + w_2 \hat{SIM} + w_3 \hat{A} + w_4 \hat{R} + w_5(1 - \hat{T}) + w_6(1 - \hat{C})
\end{equation}
\noindent with $\sum_{k=1}^{6} w_k= 1$. We call this objective function the \emph{comprehensive quality model} for service composition.
The weights can be adjusted according to users' preferences. $\hat{MT}$, $\hat{SIM}$, $\hat{A}$, $\hat{R}$, $\hat{T}$, and $\hat{C}$ are normalised values calculated within the range from 0 to 1 using Eq. (\ref{eq_normal}). To simplify the presentation we also use the notation $(Q_1,Q_2,Q_3,Q_4,Q_5,Q_6) = (MT,SIM,A,R,T,C)$. $Q_1$ and $Q_2$ have minimum value 0 and maximum value 1. The minimum and maximum value of $Q_3$, $Q_4$, $Q_5$, and $Q_6$ are calculated across all task-related candidates in the service repository $\mathcal{SR}$ using the greedy search in \cite{ma2015hybrid,da2016genetic}.

\begin{equation}
\label{eq_normal}
\hat{Q_k} = 
\begin{cases}
	\frac{Q_k - Q_{k, min}}{Q_{k, max} - Q_{k, min}} & \text{ if $k=1,\ldots,4$ and }Q_{k, max} - Q_{k, min} \neq 0,\\
	\frac{Q_{k,max} - Q_k}{Q_{k, max} - Q_{k, min}} & \text{ if $k=5,6$ and }Q_{k, max} - Q_{k, min} \neq 0,\\
	1 & \text{ otherwise}.
\end{cases}
\end{equation}

\noindent To solve the composition task satisfactorily our goal is to maximize the objective function in Eq. (\ref{eq_fitness}).

@@

\subsection{The Algorithms for our PSO-based Approach}\label{POS-based_algomargin}
The overall algorithm investigated here is made up of a PSO-based web service composition technique (Algorithm \ref{novelSteps}) and a $WG$ creating technique from a service queue (Algorithm \ref{graph_building}). In Algorithm \ref{novelSteps}, the  steps $4$, $5$, $6$ and $7$ are different from those of standard PSO: In step 4, the size of task-related service candidates generated by a greedy search determines the size of each particle's position. Each service candidate in a created service candidates queue is mapped to an index of a particle’s position vectors, where each vector has a weight value between 0.0 and 1.0. In step 5, service candidates in the queue are sorted according to their corresponding weight values in descending order. In step 6, this sorted queue is used as one of the inputs of the forward decoding Algorithm \ref{graph_building} to create a $WG$. In step 7, the fitness value of the created $WG$ is the fitness value of the particle calculated by the comprehensive model discussed in Section \ref{problemDes}.
\vspace{-0.5cm}
\begin{algorithm}
 %\LinesNumbered
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{generateWeightedGraph}{generateWeightedGraph}
 \SetKwProg{Procedure}{Procedure}{}{}
 \SetNlSty{}{}{:}
 Randomly initialise each particle in the swarm\;
  \While {max. iterations not met}{
     \ForEach{particle in the swarm}{
     Create a service candidates queue and map service candidates to a particle's position vectors\;
     Sort the service queue by position vectors' weights\;
     Use Algorithm \ref{graph_building} to create a $WG$ from the service queue\;
     Calculate the $WG$ fitness value\;

@@

\begin{algorithm}
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{createWeightedDAG}{createWeightedDAG}
 \SetKwProg{Procedure}{Procedure}{}{}
 %\LinesNumbered
 \SetNlSty{}{}{:}
 % \Procedure{}{
 \Input{ $I_T$, $O_T$, $queue$, $cache$}
 \Output{WG}
 $WG = (V, E)$\;
 $V \leftarrow$ \{$Start$, $End$ \}\;
 $OutputSet \leftarrow$ \{$I_{T}$\}\;
  \While { $O_{T}$ not satisfied by $OutputSet$}{
     \ForEach{$S$ in $queue$}{
      \uIf{$I_{S}$ satisfied by $OutputSet$}{  
        insert $S$ into $V$\;  
        adjoin $O_{S}$ to $OutputSet$\;
        $queue$.remove $S$\;   
        $e \leftarrow$ calculate $type_e$, $sim_e$ using $cache$\;
        insert $e$ into $E$\;
       }	
     }
  }
 remove $dangling$ $nodes$ and $edges$ from $WG$\; 
 \KwRet $WG$\;
 %}
 \caption{Create a $WG$ from a sorted service queue.}
\label{graph_building}
\end{algorithm}
